Comparative effectiveness of BNT162b2 versus Moderna booster doses in England: preliminary report

### 2022-05-09 write-up/report.Rmd
Update report.Rmd

@@

contrasts_overall %>%
  transmute(
    outcome_descr,
    subgroup_descr,
    subgroup_level_descr,
    n.atrisk=n.atrisk_0,
    persontime_0=label_number(1)(persontime_0/7), 
    persontime_1=label_number(1)(persontime_1/7),
    n.event_0, 
    n.event_1,
    riskCI_0 = paste0(label_number(0.01)(1-surv_0), " (", label_number(0.01)(1-surv.ul_0), "-", label_number(0.01)(1-surv.ll_0), ")"),
    riskCI_1 = paste0(label_number(0.01)(1-surv_1), " (", label_number(0.01)(1-surv.ul_1), "-", label_number(0.01)(1-surv.ll_1), ")"),
    kmrdCI = paste0(label_number(0.01)(kmrd), " (", label_number(0.01)(kmrd.ll), "-", label_number(0.01)(kmrd.ul), ")"),
    kmrrCI = paste0(label_number(0.01)(kmrr), " (", label_number(0.01)(kmrr.ll), "-", label_number(0.01)(kmrr.ul), ")"),
    irrCI =  paste0(label_number(0.01)(irr), " (", label_number(0.01)(irr.ll), "-", label_number(0.01)(irr.ul), ")"),
    #coxhrCI =  paste0(label_number(0.01)(kmrr), " (", label_number(0.01)(coxhr.ll), "-", label_number(0.01)(coxhr.ul), ")"),
  ) %>%
  group_by(outcome_descr) %>%
  mutate(
    # add indicator for row shading
    shade = cumsum((lag(subgroup_descr, 1, "")!=subgroup_descr)*1L),
    # remove repeat values of subgroup_level
    subgroup_descr = if_else(lag(subgroup_descr)==subgroup_descr & row_number()!=1, "", as.character(subgroup_descr)),
  ) %>%
  ungroup() %>%
  gt(groupname_col="outcome_descr") %>%
  cols_label(

@@

    riskCI_0 = "Risk (95% CI)",
    riskCI_1 = "Risk (95% CI)",

    kmrdCI = "Risk difference (95% CI)",
    kmrrCI = "Risk ratio (95% CI)",
    irrCI = "Incidence rate ratio (95% CI)",
    #coxhrCI = "Hazard ratio (95% CI)"
  ) %>%
  tab_spanner(
    label = "BNT162b2",
    columns = ends_with("_0")
  ) %>%
   tab_spanner(
    label = "Moderna",
    columns = ends_with("_1")
  )  %>%
   tab_spanner(
    label = "Comparison",
    columns = c("kmrrCI", "irrCI")
  ) %>%
  cols_align(
    align = c("right"),
    columns = ends_with(c("_1", "_0", "CI"))
  ) %>%
  tab_options(
    table.font.size = 8
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = shade%%2 != 0
    )
  ) %>%
  cols_hide(c("shade"))

@@

```{r, figure-km, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
plot_km <- km_estimates %>%
  filter(
    subgroup=="all",
    outcome %in% outcomes
  ) %>%
  ggplot(aes(group=treatment_descr, colour=treatment_descr, fill=treatment_descr)) +
  geom_step(aes(x=time, y=1-surv))+
  geom_rect(aes(xmin=lagtime, xmax=time, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
  geom_hline(aes(yintercept=0), colour="black")+
  facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
  scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
  scale_y_continuous(expand = expansion(mult=c(0,0.01)), )+
  coord_cartesian(xlim=c(0, NA))+
  labs(
    x="Days",
    y=NULL,
    title="Cumulative incidence",
    colour=NULL
  )+
  theme_minimal(base_size = 9)+
    theme(
      legend.position=c(0.05,0.1),
      legend.justification = c(0,0),
      legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),

@@

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.placement="outside",
      strip.text.y = element_text(angle=0),

@@

# 
# plot_kmrd <- contrasts_daily %>%
#   filter(
#     subgroup=="all",
#     outcome %in% outcomes
#   ) %>% 
#   ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
#   geom_step(aes(x=period_start, y=kmrd))+
#   geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrd.ll, ymax=kmrd.ul), alpha=0.1, colour="transparent")+
#   geom_hline(aes(yintercept=0), colour="black")+
#   facet_grid(rows=vars(outcome_descr), scales="free_y")+
#   scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
#   scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
#   scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
#   scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
#   coord_cartesian(xlim=c(0, NA))+
#   labs(
#     x="Days",
#     y=NULL,
#     title="Cumulative risk difference",
#     colour=NULL
#   )+
#   theme_minimal(base_size = 9)+
#     theme(
#       legend.position="none",
#       legend.direction = "vertical",
#       axis.text.x.top=element_text(hjust=0),
# 
#       panel.border = element_blank(),
#       axis.line.y = element_line(colour = "black"),
#       axis.line.x.top = element_line(colour = "black"),
# 
#       panel.grid.minor.x = element_blank(),
#       panel.grid.minor.y = element_blank(),
#       strip.background = element_blank(),
#       strip.text = element_blank(),
# 
# 
#       panel.spacing = unit(0.8, "lines"),
# 
#     )+
#     NULL
# 
#

plot_kmrr <- contrasts_daily %>%
  filter(
    subgroup=="all",
    outcome %in% outcomes
  ) %>%
  ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
  geom_step(aes(x=period_start, y=kmrr))+
  geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrr.ll, ymax=kmrr.ul), alpha=0.1, colour="transparent")+
  geom_hline(aes(yintercept=1), colour="black")+
  facet_grid(rows=vars(outcome_descr), scales="free_y")+
  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
  scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
  scale_y_log10(expand = expansion(mult=c(0,0.01)))+
  coord_cartesian(xlim=c(0, NA))+
  labs(
    x="Days",
    y=NULL,
    title="Cumulative risk ratio",
    colour=NULL
  )+
  theme_minimal(base_size = 9)+
    theme(
      legend.position="none",
      legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),

@@

  plot_combined <- patchwork::wrap_plots(
    plot_km, 
    #plot_kmrd, 
    plot_kmrr, 
    ncol=3
  )
  ggsave(
    filename="km.png",
    path=output_dir_rmd,
    plot=plot_combined,
  )

>>>>

Counts and survival estimates are based on values rounded up to the nearest 7, for disclosure control.

@@

contrasts_overall %>%
  transmute(
    outcome_descr,
    subgroup_descr,
    subgroup_level_descr,
    n.atrisk=n.atrisk_0,
    persontime_0=label_number(1)(persontime_0/7), 
    persontime_1=label_number(1)(persontime_1/7),
    n.event_0, 
    n.event_1,
    riskCI_0 = paste0(label_number(0.01)(risk_0), " (", label_number(0.01)(risk.ll_0), "-", label_number(0.01)(risk.ul_0), ")"),
    riskCI_1 = paste0(label_number(0.01)(risk_1), " (", label_number(0.01)(risk.ll_1), "-", label_number(0.01)(risk.ul_1), ")"),
    kmrdCI = paste0(label_number(0.01)(kmrd), " (", label_number(0.01)(kmrd.ll), "-", label_number(0.01)(kmrd.ul), ")"),
    kmrrCI = paste0(label_number(0.01)(kmrr), " (", label_number(0.01)(kmrr.ll), "-", label_number(0.01)(kmrr.ul), ")"),
    irrCI =  paste0(label_number(0.01)(irr), " (", label_number(0.01)(irr.ll), "-", label_number(0.01)(irr.ul), ")"),
    #coxhrCI =  paste0(label_number(0.01)(kmrr), " (", label_number(0.01)(coxhr.ll), "-", label_number(0.01)(coxhr.ul), ")"),
  ) %>%
  group_by(outcome_descr) %>%
  mutate(
    # add indicator for row shading
    shade = cumsum((lag(subgroup_descr, 1, "")!=subgroup_descr)*1L),
    # remove repeat values of subgroup_level
    subgroup_descr = if_else(lag(subgroup_descr)==subgroup_descr & row_number()!=1, "", as.character(subgroup_descr)),
  ) %>%
  ungroup() %>%
  gt(groupname_col="outcome_descr") %>%
  cols_label(

@@

    riskCI_0 = "10-week Risk (95% CI)",
    riskCI_1 = "10-week Risk (95% CI)",

    kmrdCI = "Risk difference (95% CI)",
    kmrrCI = "Risk ratio (95% CI)",
    irrCI = "Incidence rate ratio (95% CI)",
    #coxhrCI = "Hazard ratio (95% CI)"
  ) %>%
  tab_spanner(
    label = "BNT162b2",
    columns = ends_with("_0")
  ) %>%
   tab_spanner(
    label = "Moderna",
    columns = ends_with("_1")
  )  %>%
   tab_spanner(
    label = "Comparison",
    columns = ends_with("CI")
  ) %>%
  cols_align(
    align = c("right"),
    columns = ends_with(c("_1", "_0", "CI"))
  ) %>%
  tab_options(
    table.font.size = 8
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = shade%%2 != 0
    )
  ) %>%
  cols_hide(c("shade"))

@@

Counts and survival estimates are based on values rounded up to the nearest 7, for disclosure control.

```{r, figure-km, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
plot_km <- km_estimates %>%
  filter(
    subgroup=="all",
    outcome %in% outcomes
  ) %>%
  ggplot(aes(group=treatment_descr, colour=treatment_descr, fill=treatment_descr)) +
  geom_step(aes(x=time, y=1-surv))+
  geom_rect(aes(xmin=lagtime, xmax=time, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
  geom_hline(aes(yintercept=0), colour="black")+
  facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
  scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
  scale_y_continuous(expand = expansion(mult=c(0,0.01)), )+
  coord_cartesian(xlim=c(0, NA))+
  labs(
    x="Days",
    y=NULL,
    title="Risk",
    colour=NULL
  )+
  theme_minimal(base_size = 9)+
    theme(
      legend.position=c(0.05,0.1),
      legend.justification = c(0,0),
      legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),

@@

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.placement="outside",
      strip.text.x = element_text(angle=0),

@@

plot_kmrd <- contrasts_daily %>%
  filter(
    subgroup=="all",
    outcome %in% outcomes
  ) %>%
  ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
  geom_step(aes(x=period_start, y=kmrd))+
  geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrd.ll, ymax=kmrd.ul), alpha=0.1, colour="transparent")+
  geom_hline(aes(yintercept=0), colour="black")+
  facet_grid(rows=vars(outcome_descr), scales="free_y")+
  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
  scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
  coord_cartesian(xlim=c(0, NA))+
  labs(
    x="Days",
    y=NULL,
    title="Risk difference",
    colour=NULL
  )+
  theme_minimal(base_size = 9)+
    theme(
      legend.position="none",
      legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),

@@

plot_kmrr <- contrasts_daily %>%
  filter(
    subgroup=="all",
    outcome %in% outcomes
  ) %>%
  ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
  geom_step(aes(x=period_start, y=kmrr))+
  geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrr.ll, ymax=kmrr.ul), alpha=0.1, colour="transparent")+
  geom_hline(aes(yintercept=1), colour="black")+
  facet_grid(rows=vars(outcome_descr), scales="free_y")+
  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
  scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
  scale_y_log10(expand = expansion(mult=c(0,0.01)))+
  coord_cartesian(xlim=c(0, NA))+
  labs(
    x="Days",
    y=NULL,
    title="Risk ratio",
    colour=NULL
  )+
  theme_minimal(base_size = 9)+
    theme(
      legend.position="none",
      legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),

plot_irr <- contrasts_daily %>%
  # equivalent to the hazard
  filter(
    subgroup=="all",
    outcome %in% outcomes
  ) %>%
  ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
  geom_step(aes(x=period_start, y=irr))+
  geom_rect(aes(xmin=period_start, xmax=period_end, ymin=irr.ll, ymax=irr.ul), alpha=0.1, colour="transparent")+
  geom_hline(aes(yintercept=0), colour="black")+
  facet_grid(rows=vars(outcome_descr), scales="free_y")+
  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
  scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
  scale_y_log10(expand = expansion(mult=c(0,0.01)))+
  coord_cartesian(xlim=c(0, NA))+
  labs(
    x="Days",
    y=NULL,
    title="Hazard ratio",
    colour=NULL
  )+
  theme_minimal(base_size = 9)+
  theme(
    legend.position="none",
    legend.direction = "vertical",
    axis.text.x.top=element_text(hjust=0),

plot_combined <- patchwork::wrap_plots(
  plot_km, 
  plot_kmrd, 
  plot_kmrr, 
  #plot_irr,
  ncol=4
)

ggsave(
  filename="km.png",
  path=output_dir_rmd,
  plot=plot_combined,
)
